---
title: "Documentation"
author: "Nishat Islam"
date: "2024-03-30"
output:
  html_document: default
---
# Description
## The mars() function implements Multivariate Adaptive Regression Splines (MARS), a flexible method for regression modeling that can capture complex nonlinear relationships and interactions between variables.

# Usage
## mars(formula, data, control = mars.control())

# Arguments
## formula: A smbolic description of the model to be fitted, indicating the response variable and predictors.
## data: The dataset containing the variables mentioned in the formula.
## control: Optional. A list of parameters that control the fitting process. Default settings are applied if not specified.

# Details
## The mars() function fits a MARS model to the specified data. MARS is a non-parametric regression technique that builds models by fitting piecewise linear regressions. The function automatically identifies interaction terms and can model complex, high-dimensional datasets. The fitting process involves two main steps: a forward pass where terms are added to the model to reduce the residual sum of squares, and a backward pass where the least effective terms are pruned.

# Value
## Returns an object of class mars, which contains the fitted MARS model. This object includes components such as:

### coefficients: The model coefficients.
### terms: A list of model terms, including both basis functions and interactions.
### call: The matched call.

# Author
## Nishat Islam

# References
## Friedman, J.H. (1991). Multivariate Adaptive Regression Splines. The Annals of Statistics, 19(1), 1-67.

# See Also
## predict.mars(): Predict method for MARS models.
## summary.mars(): Summary method for MARS model objects.
## plot.mars(): Plot method for diagnostic plots of MARS model residuals.
## print.mars(): ncluding the original function call and the coefficients for each of the basis functions in the model. 
## anova.mars(): ANOVA for MARS Model Objects

# Examples
## test1 dataset of mtcars
```{r}
library(devtools)
load_all()
library(mars)
library(ISLR)
library(MASS)

# test1
data("mtcars")
mc1 = mars.control(Mmax = 10)
fit1 <- mars(mpg ~ wt + hp, data = mtcars, control=mc1)
predict1 = predict(fit1)
predict1
summary.mars(fit1)
anova.mars(fit1)
plot.mars(fit1)
print.mars(fit1)

```

## test2 dataset for Air Quality
```{r}
#test2
data("airquality")
# Clean the dataset to remove NA values for simplicity
airquality_clean <- na.omit(airquality)

mc_airquality = mars.control(Mmax = 10)
fit_airquality <- mars(Ozone ~ Solar.R + Wind + Temp, data = airquality_clean, control=mc_airquality)
predict_airquality = predict(fit_airquality)
predict_airquality
summary.mars(fit_airquality)
anova.mars(fit_airquality)
plot.mars(fit_airquality)
print.mars(fit_airquality)

```

## test3 dataset of gapminder
```{r}
#test3
# Load the gapminder dataset
if (!requireNamespace("gapminder", quietly = TRUE)) {
  install.packages("gapminder")
}
library(gapminder)
data("gapminder")

gapminder_2007 <- subset(gapminder, year == 2007)

# Setting control parameters for the MARS model
mc_gapminder <- mars.control(Mmax=10)

# Fitting the MARS model to predict life expectancy based on GDP per capita and population
fit_gapminder <- mars(lifeExp ~ gdpPercap + pop, data=gapminder_2007, control=mc_gapminder)

# Making predictions with the fitted model
predict_gapminder <- predict(fit_gapminder)

# Displaying predictions
predict_gapminder

# Summarizing the fitted model
summary.mars(fit_gapminder)

# Analyzing the model's components with ANOVA
anova.mars(fit_gapminder)

# Plotting the MARS model
plot.mars(fit_gapminder)

# Printing the model's details
print.mars(fit_gapminder)


```






